---
title: "Investigating ways of modelling data"
author: "Nikhil Parimi"
date: "2024-12-20"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(ggplot2)
library(plotly)
```

## Introduction

The Shiny App I have produced compares the difference in using a linear model as opposed to a quadratic model when the data generated is quadratic.

The formula I used to generate the synthetic data was of form $y = -0.5x^2 + 10x -18$ and in the range $x = [3, 18]$. I also generated some gaussian noise to add to the y-term of the form $y \sim \mathcal{N}(\mu =0, \sigma^2=1)$.

The code to generate this is down below:

```{r, echo = TRUE}
x_values <- seq(3, 18, by = 0.1)
y_values <- -0.5 * x_values^2 + 10 * x_values - 18
set.seed(123)
noise <- rnorm(length(y_values), mean = 0, sd = 1)
y_noisy <- y_values + noise
data <- data.frame(x = x_values, y = y_noisy)
```

## Data Visualised

Here is a visualisation of the data generated:

```{r}
gg <- ggplot(data, aes(x = x, y = y)) +
  geom_point(color = "black", size = 1.5) +
  labs(title = "Noisy Quadratic Data",
       x = "x",
       y = "y")
p <- ggplotly(gg)
p
```

## Fitting the models

In the Shiny app we then fit the two models to the data and do a visualisation of the fits as well as the models prediction for a chosen value of x (set by the user).

```{r, echo = TRUE, eval=TRUE}
# Include x^2 term in dataframe for use in quadratic model
data$x2 <- (data$x)^2
fit_linear <- lm(y~x, data = data)
fit_quad <- lm(y~x+x2, data = data)
```

## Visualising the difference between the two models

```{r}
linear_coeffs <- coef(fit_linear)
quad_coeffs <- coef(fit_quad)

linear_pred <- linear_coeffs[1] + linear_coeffs[2] * data$x
quad_pred <- quad_coeffs[1] + quad_coeffs[2] * data$x + quad_coeffs[3] * data$x2

ff <- gg +
  geom_line(aes(y = linear_pred), color = "red", linewidth = 1) +
  geom_line(aes(y = quad_pred), color = "blue", linewidth = 1) + labs(title = "Linear vs Quadratic Model Fit")

q <- ggplotly(ff)

q
```

## Residual and Diagnostic Plots of Quadratic fit
The residual plots all appears they should be for each of the diagnostic measures, showing the quadratic was a good fit!
```{r, eval = TRUE}
par(mfrow = c(2,2))
plot(fit_quad)
```